---
title: "Creating function load_vinc"
author: "Daniel Grimaldi / Schar School of Public Policy"
date: "`r Sys.time()`"
output:
  html_document:
    theme: united
    number_sections: false
---

This script creates an upper-level function (load_vinc). It loads information from the raw RAIS (unindentified) Employment-relationship files (RAIS VINC), performs basic cleaning and standardization procedures, and dump the treated data into a db file. The function will use one lower-level functions (called [**munging_rais**](file:///C:/Users/danie/OneDrive%20-%20George%20Mason%20University/03%20-%20research%20projects/Credit%20constraints%20and%20international%20trade/03%20-%20Scripts/01a-Creating-load_estab.html)) and it will accept the following arguments:

1. year: an integer value defining the year that should be loaded with information from RAIS;
2. main.source: a path to the folder in the system where RAW RAIS files are stored;
3. n.registries: the number of rows that should be read from the original raw RAIS file;
4. connection: a connection, created by dbConnect, with the db file that should receive the information;
5. table: the name of the table, inside the db.file, that should receive the information.

Since RAIS files might be too large, **load_vinc** will have no parallelism built into it.

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
if(!require(data.table)){install.packages("data.table")}
```

```{r, echo=TRUE, warning=FALSE, message=FALSE}
load_vinc <- function(year, main.source, n.registries=10, connection, table){
  
  # checking consistency for year argument
  if(!year %in% c(1994:2018)){
    stop("Please select a valid year to load RAIS data (from 1994 to 2018). The layout of the year you requested has not been incorporated to the routine yet.")
  }
  
  # Preparing extraction of raw files
  ## setting source and temporary folders for extraction
  year.source <- paste0(main.source, "/", year, ".zip")
  temp.folder <- paste0(getwd(), "/temp")
  unzip(zipfile=year.source, exdir=temp.folder)
  temp.source <- paste(temp.folder, year, sep="/")
  using.folder <- paste0(temp.source, "/using")
  dir.create(using.folder)  
  
  ## customizing for different RAIS files structures
  if(year==2018){
    folder <- grep("using", list.dirs(temp.source, recursive = FALSE), value = TRUE, invert=TRUE)
    vinc.files <- list.files(folder, full.names =TRUE)   
  }else{
    vinc.files <- paste(temp.source,
                        grep("^[A-Z][A-Z][0-9][0-9][0-9][0-9]", list.files(temp.source), value = TRUE),
                        sep="/")
  }
  
  # defining sectorial classification
  sector.class <- ifelse(year<=2001, "cnae", "cnae10")
  
  ## extracting and defining files to load
  command.vinc.i <- paste("7z x ", shQuote(vinc.files), " -o", shQuote(using.folder), "", sep="")
  lapply(command.vinc.i, system)
  file.i <- list.files(using.folder, full.names=TRUE)
  
  # Loading files and munging raw information before dumping it into db file
  ## preparing counting for the creation of unique key for employment registry
  count.start=1
  ## looping files and dumping info into data lake
  for (i in file.i){
    data.i <- fread(i, nrows=n.registries)
    ## basic munging and stardadization 
    data.i <- munging_rais(data.i, year, type="VINC") %>% 
    ## customizing to fit rp2.db
    filter(wage.year>0) %>%
      rename(employment_wage_year = wage.year,
             employment_wage_dec = wage.dec,
             employment_active_dec = active.contract,
             employment_hours = hours.hired,
             employment_duration = time.employment,
             employment_type = type.contract,
             worker_age = worker.age,
             worker_nationality = worker.nationality,
             worker_education = worker.education,
             worker_gender = worker.gender,
             company_type = type.company,
             company_legal_status_1994 = legal.status.1994,
             company_legal_status = legal.status.pos1994,
             mun_id = mun.id, 
             rais_sector_id = cnae95) %>%
      mutate(mun_id = as.character(mun_id),
             rais_sector_classification = sector.class) %>%
      select(year, employment_wage_year, employment_wage_dec, employment_active_dec, employment_hours, employment_duration, employment_type, worker_age, worker_nationality, worker_education, worker_gender, company_type, company_legal_status_1994, company_legal_status, mun_id, rais_sector_id, rais_sector_classification)
    data.i$employment_record_id  <- paste("e", year, count.start:(count.start+nrow(data.i)-1), sep="-")
    data.i <- relocate(data.i, employment_record_id)
    # set new start for next file (keys must be consistenti within years)
    count.start <- count.start+nrow(data.i) 
    
  # write info into data lake
  dbWriteTable(connection, table, data.i, append=TRUE)
  print(paste0(i, " dumped into db file."))
  
  # remove data and clean memory before loading another file
  rm(data.i)
  gc(reset=TRUE)
  }
  
  # removing temporary files from memory
  unlink(file.i, recursive = FALSE)
  unlink(temp.folder, recursive = TRUE)
  
  print(paste0("RAIS VINC for year ", year, " dumped into db file."))
}
dump("load_vinc", file="load_vinc.R")
```