---
title: "Creating the sectorial crosswalk file (Original RAIS sector codes to standardized economic sectors)"
author: "Daniel Grimaldi / Schar School of Public Policy"
date: "`r Sys.time()`"
output:
  html_document:
    theme: united
    number_sections: false
bibliography: references.bib
---

```{r, warning=FALSE, message=FALSE}
library(knitr)
if(!require(kableExtra)){install.packages("kableExtra")}
if(!require(tidyverse)){install.packages("tidyverse")}
if(!require(magrittr)){install.packages("magrittr")}
if(!require(readxl)){install.packages("readxl")}
if(!require(data.table)){install.packages("data.table")}
source("munging_rais.R", encoding = "Windows-1252")
source("load_estab.R")
```

# Introduction

In the RAIS raw files, information on sectorial classification reported in the CNAE-1995 field pertains to two different classification. Until 2001, codes are reported according to CNAE-1995. From 2002 onward, codes are reported according to a revised version of CNAE-1995 (called CNAE 1.0). There are some relevant changes between these two systems. CNAE 1.0 has a lager number of sectors -- that is, it brings a more disaggregated view of the economy. In order to keep time-comparability while minimizing the number of multiple associations (cases when 1 code from one system has to be splitted among distinct codes in the other system), this script will build a crosswalk file allowing researchers to move from CNAE 1.0 to CNAE. This crosswalk used information from IBGE with only smaller adaptions^[Information about CNAE-CNAE 1.0 compatibility can be obtained at https://concla.ibge.gov.br/images/concla/documentacao/CNAE1.0xCNAE.pdf]. This information wwas used to manually implement a series of associations between CNAE 1.0 codes and CNAE codes^[The file that consolidates this series of associations can be consulted at [manual_imputation_cnae10_cnae.csv]( https://exchangelabsgmu-my.sharepoint.com/:x:/g/personal/dgrimald_masonlive_gmu_edu/EYmKW3ufFbpMqqXQxydE_VgBl7M97fFv2u7rvIoxqCy1tg?e=fwa309).] 

After that, the main challenge was to build a splitting rule for the cases with multiple associations. To avoid inconsistencies that could jeopardize the causal identification desgined in @grimaldi2021, this code will use the average economic structure between years 1999-2001 (the last three before the adoption of CNAE 1.0 by RAIS) to build two different splitting rules. We will use an example to explain these rules. Consider that sector $X$ in CNAE 1.0 is associated with sectors $X1$ and $X2$ in CNAE. Also, by analyzing data from RAIS raw files from years 1999-2001, assume that $X1$ employees on average 60 workers per year, while $X2$ is associated with 40 workers per year. The first splitting rule will direct $X$ employees, as reported from 2002 onward, integrally to sector $X1$ because it is the main CNAE sector associated with sector $X$. Considering the same example, the second rule will divide the total amount of employees reported to sector $X$ in the following manner: 60% to $X1$ and 40% to $X2$.

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
# downloading file with CNAE information from IBGE
temp <- tempfile(fileext = ".xls")
download.file(url="https://cnae.ibge.gov.br/images/concla/downloads/CNAE.xls",
              destfile = temp,
              method="wininet",
              mode="wb")

# munging data on CNAE sectors
cnae.sectors<- read_excel(temp, skip=5)
cnae.sectors <- cnae.sectors[,c(4:5)]
names(cnae.sectors) <- c("cnae_id", "cnae_description")
cnae.sectors %<>%
  mutate(cnae_id = gsub("[:.:]", "", cnae_id),
         cnae_id = as.numeric(gsub("-", "", cnae_id)),
         cnae_version = "cnae") %>%
  select(cnae_id, cnae_description, cnae_version) %>% 
  na.omit() %>% 
  as.data.frame()

# downloading file with CNAE v1.0 information from IBGE
temp <- tempfile(fileext = ".xls")
download.file(url="https://cnae.ibge.gov.br/images/concla/downloads/CNAE1.0.xls",
              destfile = temp,
              method="wininet",
              mode="wb")

# munging data on CNAE sectors
cnae10.sectors<- read_excel(temp, skip=5)
cnae10.sectors <- cnae10.sectors[,c(4:5)]
names(cnae10.sectors) <- c("cnae_id", "cnae_description")
cnae10.sectors %<>%
  mutate(cnae_id = gsub("[:.:]", "", cnae_id),
         cnae_id = as.numeric(gsub("-", "", cnae_id)),
         cnae_version = "cnae10") %>%
  select(cnae_id, cnae_description, cnae_version) %>% 
  na.omit() %>% 
  as.data.frame()

# loading manually imputed relations to change CNAE 1.0 into CNAE (now called sector_id)
cnaes.relations <- read.csv("../02 - Data sets/manual_imputation_cnae10_cnae.csv", encoding = "UTF-8")
cnaes.relations <- cnaes.relations[,c(1,2)]
names(cnaes.relations) <- c("cnae10_id", "cnae_id")

# loading RAIS Company files for the 1999-2001 period
rais.data <- rbindlist(lapply(c(1999, 2000, 2001),
                              load_estab,
                              main.source = "C:/Users/danie/OneDrive - George Mason University/01 - datalake/RAIS/RAIS - LAI",
                              output="Rdata",
                              n.registries=Inf))

rais.data %<>%
  select(sector_rais_id, total_emp) %>% 
  group_by(sector_rais_id) %>% 
  summarise(total_emp_cnae = sum(total_emp))

# calculating the weights (considering relative employment importance between 1999-2001 years) for each CNAE1.0 - CNAE relation
crosswalk1 <- cnaes.relations %>% 
  left_join(rais.data, by=c("cnae_id"="sector_rais_id")) %>% 
  group_by(cnae10_id) %>%
  mutate(cnae_version="cnae10",
         n.cnaes = n_distinct(cnae_id),
         total_emp_cnae10 = sum(total_emp_cnae, na.rm = TRUE),
         max_emp_cnae = max(total_emp_cnae, na.rm=TRUE),
         v1 = case_when(n.cnaes==1 ~ 1,
                        n.cnaes>1 & is.na(total_emp_cnae) ~ 0,
                        TRUE ~ as.numeric(total_emp_cnae==max_emp_cnae)),
         v2 = case_when(n.cnaes==1 ~ 1,
                        n.cnaes>1 & is.na(total_emp_cnae) ~ 0,
                        TRUE ~ total_emp_cnae/total_emp_cnae10)) %>% 
  ungroup() %>% 
  select(cnae10_id, cnae_version, cnae_id, v1, v2) %>% 
  rename(rais_sector = cnae10_id,
         sector_id = cnae_id)

# The weights for the CNAE - CNAE crosswalk are all equal to 1
crosswalk2 <- cnae.sectors %>% 
  mutate(cnae_version="cnae",
         sector_id=cnae_id,
         v1=1,
         v2=1) %>%
  rename(rais_sector = cnae_id) %>% 
  select(rais_sector, cnae_version, sector_id, v1, v2)

# consolidating the crosswalk
rais_standard.crosswalk <- rbind(crosswalk1, crosswalk2)

# saving the crosswalk
save(rais_standard.crosswalk, file="../02 - Data sets/rais-standard_crosswalk.Rdata")
```








